---
title: "Hotel Bookings"
author: "Katie Romrell"
date: "2025-04-20"
output:  html_document
---
# Context and Set-up

This is a walkthrough of the analysis I've completed with a sample dataset about hotel bookings given during the Google Data Analyst Certification. In this hypothetical situation (invented by me, not the course), the hotel is planning to do an ad campaign targeted toward families since family reservations generate more revenue than single reservations. That said, the hospital executives are concerned that bookings with families might have disproportionately high complications (e.g., more cancellations, less lead-time (advanced bookings), more special requests), so they want to explore any current relationships before rolling out the campaign.

Analysis Questions:
1. What is are the current rates of family reservations and complications separate from each other?
2. What, if any, is the relationship between hotel stay complications and family bookings?

## Setting up the Environment: Loading packages and dataset.
```{r install packages}
install.packages("tidyverse")
library(tidyverse)
install.packages("skimr")
library(skimr)
install.packages("janitor")
library(janitor)

hotel_bookings <- read_csv("C:/Users/kmrom/Downloads/hotel_bookings.csv")
```

## Preview Dataset: exploring the data

```{r explore data}
View(hotel_bookings)
colnames(hotel_bookings)
skim_without_charts(hotel_bookings)
clean_names(hotel_bookings)

```
## Clean Data
In order to make my future analysis easier, I cleaned the data in the following ways:

**1. Deleted Outliers and Observations with NA:** Out of 120,000 observations, there were only 4 NA values and 3 instances with more than three children (two 10 and one 9). These values were removed to make the scale of charts more streamlined (since they were statistically negligible). 

**2. Create New Columns:** I created an all_children column that combines the pre-existent children and babies columns. I also created a text-based column identifying if the booking has "Children" or "No Children" to help with future bar graphs. Additionally, in (hypothetically) talking to the hotel staff, they said that lead-time was only experienced as a complication if it was under 7 days, so I created a binary column called "lead_under_7". 

**4. Renamed Column:** I simplified "total_of_special_requests" as "special_request"

```{r clean data}
hotel_bookings_clean<-hotel_bookings %>% 
  drop_na(children) %>% 
  mutate(all_children=children+babies) %>% 
  mutate(presence_of_children=ifelse(children>0, "Children", "No Children")) %>% 
  mutate(presence_of_children_binary=ifelse(children>0, 1, 0)) %>% 
  mutate(lead_under_7=ifelse(lead_time<7, 1, 0)) %>% 
  filter(all_children<9) %>% 
  rename(special_requests=total_of_special_requests)
skim_without_charts(hotel_bookings_clean)
View(hotel_bookings_clean)

```
# Analysis
Below is my analysis. As a reminder, the analysis priorities are to 1) understand the current rate of family reservations and complications, and 2) to understand the relationship between children and reservation complications such as cancellations, lack of lead-time, and special requests. 

## Descriptive Statistics
For this analysis, I want to focus on factors that complicate a hotel stay, such as cancellations, lack of lead-time, and specials requests.To do this, I'll start by aggregating the data to get a feel for the norms. I included both mean and median for lead_time because I was worried the distribution may be skewed.

```{r aggregate data}
hotel_bookings_clean %>% 
  summarize(
    cancel_perc=mean(is_canceled)*100,
    lead_under_7_tally=sum(lead_under_7),
    requests_mean=mean(special_requests),
    requests_max=max(special_requests),
    requests_min=min(special_requests)
  )

hotel_bookings_clean %>% 
  summarize(
    children_mean=mean(all_children),
    reservations_with_children=sum(presence_of_children_binary),
    total_reservations=length(presence_of_children_binary),
    perc_with_children=sum(presence_of_children_binary)/length(presence_of_children_binary)*100
  )


```

## Plot Relationships
Now that I have a better idea of the norms, I want to plot some of these relationships. These aren't charts I would show to stakeholders, so I didn't prioritize aesthetics in the chart.

```{r plot data}
ggplot(data=hotel_bookings_clean)+
  geom_bar(mapping=aes(x=is_canceled, fill=presence_of_children))

ggplot(data=hotel_bookings_clean)+
  geom_bar(mapping=aes(x=special_requests, fill=presence_of_children))

hotel_bookings_clean %>% 
  filter(lead_under_7==1) %>% 
  ggplot()+geom_jitter(mapping=aes(x=lead_time, y=all_children), alpha=.5)

```
Because of the extreme differences in the n values of those with and without kids, it's difficult for these visualizations to show any relationship that may exist. For that reason, I decided to run new vizes on the averages.

```{r}
lead_time_averages<-hotel_bookings_clean %>% 
  group_by(all_children) %>% 
  summarize(
    mean=mean(lead_time),
    
    )
  

lead_time_averages
```

